---
title: "Amundsen 2021 - Analyses"
author: "Pierre Priou"
date: "17/01/2022"
output: github_document
---

```{r r-setup, message = FALSE}
# Load packages
library(tidyverse)                # Tidy code
library(lubridate)                # Date handling
library(kableExtra)               # Pretty tables
library(cowplot)                  # Multiple plots
library(png)                      # Read png files
library(rgdal)                    # Read shapefiles
library(marmap)                   # Bathymetry data of the Arctic Ocean
library(scales)                   # Plots with dates
library(stringr)                  # Work with strings
library(MetBrewer)                # Colour palettes
library(cmocean)                  # Colour palettes
source("R/sv_correction_fm.R")    # Correct Sv vertical profiles
# Suppress summarise() warning
options(dplyr.summarise.inform = F)
# Theme for figures
theme_set(theme_bw())
theme_update(panel.grid = element_blank(), 
             panel.border = element_blank(), 
             axis.line = element_line(),
             axis.text = element_text(size = 10),
             axis.title = element_text(size = 11),
             legend.title = element_text(size = 11),
             strip.text.x = element_text(size = 11, face = "plain", hjust = 0.5),
             strip.background = element_rect(colour = "transparent", fill = "transparent"),
             legend.margin = margin(0,0,0,0),
             legend.box.margin = margin(0,0,-8,0),
             plot.margin = unit(c(0.05,0.05,0.05,0.05), "in"))
```

# Introduction

Diel vertical migrations (DVM) from ocean's depth to the surface are a ubiquitous pattern of pelagic organisms of the global ocean (Cohen 2009). While DVM are ultimately driven by predator-prey interactions (e.g., Urmy et al. 2021), environmental factors are known to define the extent and timing of these migrations. For example, temperature and oxygen can influence DVM patterns (Bianchi et al. 2013, Benoit-Bird et al. 2015?), but light penetration is considered to trump those other environmental cues (Asknes et al. 2017). Fish living in the mesopelagic zone, where the light intensity is between 10^-9^ to 10^-1^ μmol quanta m^2^ s^-1^ (Kaartvedt et al., 2019), track a narrow range of ambient light levels—their light comfort zone (Rostad 2016)—and migrate vertically following those isolumes during the day-night cycle. They track their antipredation window, which enhance their feeding chances while reducing mortality through predation (Clark & Levy, 1988). 

In the Arctic, the extreme light regime has the potential to disrupt traditional DVM patterns, which could prevent the establishment of viable mesopelagic fish populations in the Arctic (Kaartvedt 2008). Yet, DVM can occur during the polar night at low light levels (Berge et al. 2009) and mesopelagic fish are not as seldom in the Arctic as previously thought, and are, in fact, widespread in high Arctic Seas (Priou et al. *in prep*). However, their survival mechanisms remain elusive. For most of the year mesopelagic fish remain segregated from their favourite preys (Gjøsæter et al. 2017, Knutsen et al. 2017, Geoffroy et al. 2019, Priou et al. 2021). Autumn has been suggested to be a key period for feeding of mesopelagic fish, as the photoperiod alternates between day and night---favouring large scale DVM---and their copepod preys occupy most of the water column (**ref**).

The goals of this study were two-fold, (1) investigate whether mesopelagic organisms conduct DVM in autumn, and (2) if those organisms were associated to a specific range of ambient light levels matching those of the global ocean. Here, we deployed a combination of acoustic, optical, and environmental sensors and nets to document the vertical distribution of pelagic fish and zooplankton in northern Baffin Bay in autumn. We test the hypothesis that pelagic organisms follow a narrow range of light levels (different from the "new" definition of the mesopelagic zone) and thus conduct DVM.

# Materials and methods

```{r load-data, message = FALSE}
# Load data
load("data/CTD/DarkEdge_CTD.RData") # CTD data
bathy <- read_csv("data/bathy/bathy_BaffinBay_-105-45;65;82_res2.csv") %>%
  mutate(depth_d = factor(if_else(between(depth,-50,Inf), "0-50",
                        if_else(between(depth,-100,-50), "50-100",
                        if_else(between(depth,-200,-100), "100-200",
                        if_else(between(depth,-500,-200), "200-500",
                        if_else(between(depth,-1000,-500), "500-1000",
                        if_else(between(depth,-Inf,-1000), "1000-2500", ">0")))))),
                 levels = c(">0","0-50","50-100","100-200","200-500","500-1000","1000-2500")))
coast <- readOGR("data/bathy/ne_10m_land.shp", verbose = F) %>% # Coastlines
  fortify() %>%
  rename(lon = long,
         region = id)
glacier <- readOGR("data/bathy/ne_10m_glaciated_areas.shp", verbose = F) %>% # Glaciers
  fortify() %>%
  rename(lon = long,
         region = id)
```
