---
title: "Extrapolation of irradiance using C-OPS and TriOS data"
author: "Pierre Priou - pierre.priou@mi.mun.ca"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
---

```{r load-packages, message=FALSE}
library(tidyverse)    # Tidy coding
library(lubridate)    # Dates formatting
library(plotly)       # Interactive plots
library(cowplot)      # Panel plots
library(photobiology) # Associate wavelength to colour
library(dtplyr)       # Speed-up calculations
library(broom)        # Tidy linear model
```

```{r figure-theme}
theme_set(theme_bw())
theme_update(panel.border = element_blank(), 
             axis.line = element_line(),
             axis.text = element_text(size = 10),
             axis.title = element_text(size = 11),
             legend.title = element_text(size = 11),
             strip.text.x = element_text(size = 11, face = "plain", 
                                         hjust = 0.5),
             strip.background = element_rect(colour = "transparent", 
                                             fill = "transparent"),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             plot.margin = unit(c(0.075, 0.075, 0.075, 0.075), "in"))
```

# Aim

This document uses data from the Compact Optical Profiling System (C-OPS) and TriOS surface irradiance sensor to estimate underwater irradiance at mesopelagic depth throughout the DarkEdge campaign. To achieve this we need to: 

1. Convert C-OPS data from energy units (μW cm^-2^ nm^-1^) to quantum flux (μmol photons m^-2^ s^-1^ nm^-1^). 
2. Use the attenuation coefficient (K~d~) from C-OPS data to calculate underwater irradiance from TriOS data.
3. Evaluate the quality of the extrapolation by comparing observed underwater irradiance from C-OPS with the calculated one from TriOS.

# C-OPS quantum flux and K~d~ estimation

For our analysis we selected wavelength from 400 to 690 nm as the data quality of higher wavelength was not optimal. 

```{r COPS-loading}
COPS_raw <- read_csv("data/C-OPS/C-OPS_raw_processed.csv", show_col_types = F,
                     locale = locale(tz = "UTC")) %>%
  # Add metadata
  mutate(day = day(date_COPS),
         month = month(date_COPS))
```

I selected one replicate of each C-OPS cast at each station.

I use the following equation to calculate the quantum flux (E~d~ in μmol photons m^-2^ s^-1^ nm^-1^) from the irradiance (I in W m^-2^ nm^-1^:

$$E{d} = I \times \lambda \times 0.836 \times 10^{-2}$$

where E~d~ is the downwelling irradiance in μmol photons m^-2^ s^-1^ nm^-1^ and $\lambda$ is the wavelength in nm.

```{r COPS-quantum-flux-kd-calculation}
COPS <- COPS_raw %>%
  # Select wavelength with less variation
  filter(between(lambda_nm, 400, 690)) %>%
  # Select casts
  filter(COPS_ID %in% c("DE110_3", "DE120_2", "DE130_2", "DE320_3", "DE410_10",
                        "DE420_11", "DE430_2", "DE440_2")) %>%
  # Remove NAs
  filter(is.na(ed_uW_cm2) == F) %>%
  # Unique ID for each profile/wavelength for plotting each profile
  unite(ID, COPS_ID, lambda_nm, sep = "_", remove = F) %>%
  # Calculate quantum flux
  mutate(ed_W_m2 = ed_uW_cm2 * 10^-2,
         ed_umol_m2 = ed_W_m2 * lambda_nm * 0.836 * 10^-2)
```

ln(E~d~) decrease linearly throughout the water column so Kd can be estimated using z~1~ close to the surface and z~2~ close to the deepest measurement. 

```{r plot-Ed-station, warning=FALSE}
# Associate each wavelength to appropriate color
col_pal <- COPS %>%
  group_by(lambda_nm) %>%
  summarise() %>%
  as.matrix() %>%
  # Find color for each wavelength
  w_length2rgb() %>%
  as_tibble() %>%
  as.matrix()

ggplotly(
  COPS %>%
    # Select cast to facilitate display
    filter(COPS_ID %in% c("DE110_3", "DE120_2", "DE130_2", "DE320_3", "DE410_10",
                          "DE420_11", "DE430_2", "DE440_2")) %>%
    mutate(day = paste0(day, "/10 - ", station)) %>%
    ggplot() +
    geom_path(aes(x = ed_umol_m2, y = depth, col = factor(lambda_nm), group = ID),
              alpha = 0.5, na.rm = T) +
    scale_x_log10("log10(Ed)",
                  breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
                  labels = scales::trans_format("log10", scales::math_format(10 ^ .x))) +
    scale_y_reverse("Depth (m)") +
    scale_color_manual("Wavelength (nm)", values = col_pal) +
    facet_wrap(~day, scales = "fixed") 
)
```

To estimate K~d~ we used a linear regression of ln-transformed E~d~ over depth in 5 m depth cell. K~d~ is equal to the inverse of the slope estimate: 

$$K{d}(\lambda) = -\frac{ln({\frac{ E{d}(z_{2}, \lambda)} {{E{d}(z_{1}, \lambda)}})}} {(z_{2} - z_{1})}$$

where K~d~ is the attenuation coefficient in m^-1^, E~d~ is the downwelling irradiance (in μmol photons m^-2^ s^-1^ nm^-1^) at depth z~1~ (closer to the surface) and z~2~ (deeper down).

We remove K~d~ with a coefficient of determination (R^2^) below 0.9. 

```{r Kd-estimation}
Kd <- COPS %>%
  # Round depth
  mutate(depth_round = plyr::round_any(depth, 5, floor)) %>%
  # Adjust variable class for model
  mutate(day = factor(day),
         station = factor(station),
         lambda_nm = factor(lambda_nm),
         depth_round = factor(depth_round)) %>%
  # Nest data
  nest(data = c(-day, -station, -lambda_nm, -depth_round)) %>%
  # Fit linear model for each day, wavelength, and 10 m depth bin
  mutate(fit = map(data, ~ lm(log(ed_umol_m2) ~ depth, data = .x)),
         tidied = map(fit, tidy),
         glanced = map(fit, glance)) %>%
  # Remove unused variables
  select(-data, -fit) %>%
  # Unnest and extract R2, Kd estimate, and number of observations
  unnest(tidied) %>%
  filter(term == "depth") %>%
  select(-std.error, -statistic, -p.value) %>%
  unnest(glanced) %>%
  select(-term, -adj.r.squared, -sigma, -statistic, -p.value, -df, -logLik, -AIC, -BIC,
         -deviance, -df.residual) %>%
  # Rename data
  rename(R2 = r.squared,
         kd = estimate,
         n = nobs) %>%
  mutate(kd = kd * -1) %>%
  # Select data with R2 > 0.9 and n > 3
  filter(R2 > 0.9 & n > 3) %>%
  # Correct variable class
  mutate(lambda_nm = as.numeric(as.character(lambda_nm)),
         day = as.integer(as.character(day)), 
         depth_round = as.numeric(as.character(depth_round))) %>%
  # Rename variables to match TriOS data
  rename(depth = depth_round)
```

K~d~ estimates seemed realistic. 

```{r plot-Kd}
# Associate each wavelength to appropriate color
col_pal <- Kd %>%
  group_by(lambda_nm) %>%
  summarise() %>%
  as.matrix() %>%
  # Find color for each wavelength
  w_length2rgb() %>%
  as_tibble() %>%
  as.matrix()

ggplotly(Kd %>%
           mutate(day = paste0(day, "/10 - ", station)) %>%
           ggplot() +
           geom_path(aes(x = kd, y = depth, col = factor(lambda_nm)),
                     alpha = 0.5, na.rm = T) +
           scale_y_reverse("Depth (m)") +
           scale_color_manual("Wavelength (nm)", values = col_pal) +
           guides(color = guide_legend(ncol = 2)) + 
           facet_wrap(~day, scales = "free_y") 
)
```

To extrapolate light et depth below the last C-OPS measurement I used the deepest K~d~ available for each wavelength. 

```{r Kd-max-depth}
Kd_max <- Kd %>%
  group_by(day, station, lambda_nm) %>%
  # Find Kd at max depth
  mutate(max_depth = if_else(is.na(kd) == F, depth, NaN),
         max_kd_depth = max(max_depth, na.rm = T),
         keep = if_else(max_depth == max_kd_depth, T, F)) %>%
  ungroup() %>%
  filter(keep == T) %>%
  # Rename and keep variables of interest
  rename(kd_max_depth = kd) %>%
  select(day, lambda_nm, kd_max_depth)
```


# Underwater irradiance from TriOS data

**Important, October 15th may not have been processed as it contains outliers (negative irradiances and extremely high values > 1 x 10^8^). For now I have excluded this date from the final dataset.** 

```{r TriOS-loading}
trios_raw <- read_csv("data/TriOS/TriOS_processed.csv", show_col_types = F,
                     locale = locale(tz = "UTC")) %>%
  mutate(day = day(date_trios),
         month = month(date_trios)) %>%
  # Remove outliers from October 15th
  filter(trios_filename != "15Oct") %>%
  # Remove unused variables
  select(-trios_filename, -lat, -lon) %>%
  # Select wavelength that match those of the C-OPS
  filter(between(lambda_nm, 400, 690))
```

The TriOS surface irradiance sensor recorded data at 1 sec interval and 1 nm interval from 320 to 950 nm. To handle this huge dataset, I selected only the wavelength which matched those to the C-OPS, and calculated the mean surface irradiance in 5 min intervals. Underwater irradiances profiles is then extrapolated from surface down to the seafloor (or deepest CTD measurement) from this equation:

$$E_{d}(z_{2}, \lambda) = 0.97 \times E_{0}(\lambda) \times e^{(-K_{d}(\lambda) \times (z_{2} - z_{1}))}$$

where E~d~(z~2~, $\lambda$) is the underwater irradiance of wavelength $\lambda$ at depth z2, 0.97 is a constant for reflectance of irradiance at the water surface, E~0~($\lambda$) is the irradiance of wavelength $\lambda$ at the surface, K~d~($\lambda$) is the attenuation coefficient of wavelength $\lambda$, and z~1~ and z~2~ is the depth interval. The depth interval is here set to 5 m to match the vertical resolution of C-OPS data.

```{r extrapolation-underwater-irradiance}
# Set depth of interpolation
ed_calc_umol_m2 <- trios_raw %>% 
  # Add depth for completing dataset
  mutate(depth = 0,
         day = day(date_trios)) %>%
  # Select days that have a Kd
  filter(day %in% c(12, 13, 14, 17, 20, 21, 22, 23)) %>%
  # Call dtplyr to speed-up calculations
  lazy_dt() %>% 
  # Calculate average E0 every 5 min
  mutate(date_round = round_date(date_trios, "5 mins")) %>%
  group_by(date_round, day, lambda_nm, depth) %>%
  summarise(e0_umol_m2 = mean(e0_umol_m2, na.rm = T)) %>%
  ungroup() %>%
  # Complete dataset by adding depth
  group_by(date_round, day, lambda_nm, e0_umol_m2) %>%
  complete(depth = seq(0, 1500, 5)) %>%
  ungroup() %>%
  # Add Kd
  left_join(., Kd, by = c("day", "lambda_nm", "depth")) %>%
  # Add Kd at max  depth
  left_join(., Kd_max, by = c("day", "lambda_nm")) %>%
  group_by(day, lambda_nm, e0_umol_m2) %>%
  mutate(kd_composite = if_else(is.na(kd) == F, kd, kd_max_depth)) %>%
  ungroup() %>%
  filter(is.na(kd_max_depth) == F) %>% 
  # Calculate downwelling irradiance
  mutate(ed_calc_umol_m2 = 0.97 * e0_umol_m2 * exp(- kd_composite * depth)) %>%
  as_tibble() %>%
  # Select relevant variables
  select(date_round, day, lambda_nm, depth, kd_composite, ed_calc_umol_m2)
```

Plot data.

```{r plot-ed-490nm, warning=FALSE, fig.width=8, fig.height=6}
plot_grid(
  ed_calc_umol_m2 %>%
    filter(lambda_nm == 490 & between(day, 20, 21) & depth < 600) %>%
    mutate(alpha = if_else(between(log10(ed_calc_umol_m2), -9, -1), "in", "out")) %>%
    ggplot() +
    geom_tile(aes(x = date_round, y = depth, fill = log10(ed_calc_umol_m2), alpha = alpha)) +
    geom_contour(aes(x = date_round, y = depth, z = log10(ed_calc_umol_m2)),
                 binwidth = 2, col = "white") +
    scale_fill_viridis_c("log10(Ed)\n490nm", option = "mako") +
    scale_x_datetime(breaks = seq(ymd_hms("2021-10-20 00:00:00"), ymd_hms("2021-10-21 18:00:00"), "6 hours"),
                     date_labels = "%d/%m\n%H:%M") +
    scale_alpha_discrete(range = c(2, 0.2)) +
    scale_y_reverse("Depth (m)") + 
    guides(alpha = "none") +
    ggtitle(expression("Kaartvedt et al. 2019 – 10"^-1*" 10"^-9*"")) +
    theme(axis.title.x = element_blank()),
  ed_calc_umol_m2 %>%
    filter(lambda_nm == 490 & between(day, 20, 21) & depth < 600) %>%
    mutate(alpha = if_else(between(log10(ed_calc_umol_m2), -9, -6), "in", "out")) %>%
    ggplot() +
    geom_tile(aes(x = date_round, y = depth, fill = log10(ed_calc_umol_m2), alpha = alpha)) +
    geom_contour(aes(x = date_round, y = depth, z = log10(ed_calc_umol_m2)),
                 binwidth = 2, col = "white") +
    scale_fill_viridis_c("log10(Ed)\n490nm", option = "mako") +
    scale_x_datetime(breaks = seq(ymd_hms("2021-10-20 00:00:00"), ymd_hms("2021-10-21 18:00:00"), "6 hours"),
                     date_labels = "%d/%m\n%H:%M") +
    scale_alpha_discrete(range = c(2, 0.2)) +
    scale_y_reverse("Depth (m)") + 
    guides(alpha = "none") +
    ggtitle(expression("Asknes et al. 2017 – 10"^-6*" 10"^-9*"")) +
    theme(axis.title.x = element_blank()),
  ed_calc_umol_m2 %>%
    filter(lambda_nm == 490 & between(day, 20, 21) & depth < 600) %>%
    mutate(alpha = if_else(between(ed_calc_umol_m2, 2.5 * 10^-7, 1.6 * 10^-5), "in", "out")) %>%
    ggplot() +
    geom_tile(aes(x = date_round, y = depth, fill = log10(ed_calc_umol_m2), alpha = alpha)) +
    geom_contour(aes(x = date_round, y = depth, z = log10(ed_calc_umol_m2)),
                 binwidth = 2, col = "white") +
    scale_fill_viridis_c("log10(Ed)\n490nm", option = "mako") +
    scale_x_datetime(breaks = seq(ymd_hms("2021-10-20 00:00:00"), ymd_hms("2021-10-21 18:00:00"), "6 hours"),
                     date_labels = "%d/%m\n%H:%M") +
    scale_alpha_discrete(range = c(2, 0.2)) +
    scale_y_reverse("Depth (m)") + 
    guides(alpha = "none") +
    ggtitle(expression("Norheim et al. 2016 – 1.6 x 10"^-5*" 2.5 x 10"^-7*"")) +
    theme(axis.title.x = element_blank()),
  ncol = 1, labels = "AUTO", align = "hv", axis = "tblr")
```

# Save data

Save extrapolated values. 

```{r extrapolation }
write_csv(ed_calc_umol_m2, file = "data/TriOS/Ed_extrapolated.csv")
```

